An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-03-26 EconGrowthSeminar.Rnw
Now including "comments" frames about each model.

@@

@@

\begin{frame}{Parameter Estimation for Cobb-Douglas With Energy} %----------------------------
  \begin{center}\cdeEqn\end{center}
  \begin{itemize}
    \item{Objective: determine $\lambda$, $\alpha$, $\beta$, and $\gamma$ that best fits historical data}
    \item{Almost identical to Cobb-Douglas without energy}
    \item{Perform grid search over $\gamma$}
    \item{Re-estimate $\lambda$, $\alpha$, and $\beta$ at each value of $\gamma$}
    \item{Respect sum constraint: $\alpha + \beta + \gamma = 1$}
    \item{Respect absolute constraints: $0 \leq \alpha, \beta, \gamma \leq 1$} (not yet)
  \end{itemize}
\end{frame}

@@

@@

@@

@@

@@

\begin{frame}{Example LINEX Fit} %----------------------------
  Heun complete
\end{frame}

@@

\begin{frame}{Tentative Conclusions} %----------------------------
  \begin{itemize}
  \item{Including energy is unwarranted \\
        \scriptsize
        Little statistical evidence that including energy in production functions improves stastical properties of fitting parameters.
        \normalsize}
  \item{Including energy is unnecessary \\
        \scriptsize
        Energy is \emph{always already} embedded in capital stock ($K$) and labor ($L$).
        \normalsize}
  \vspace{\baselineskip}
  \item{Need a modeling method that captures energy's embeddedness \\
        \scriptsize
        System dynamics modeling
        \normalsize}
  \end{itemize}
\end{frame}

\begin{frame}{Next Steps (Summer 2013)} %----------------------------
  \begin{itemize}
    \item{Estimate statistical properties for \emph{all} fitting parameters}
    \vspace{\baselineskip}
    \item{Incorporate all constraints for Cobb-Douglas model}
    \vspace{\baselineskip}
    \item{Complete paper}
  \end{itemize}  
\end{frame}

>>>>

\begin{frame}{Comments on Single-factor Model Equation} %----------------------------
    \begin{center} \sfEqn \end{center}
    \vspace{\baselineskip}

    \begin{itemize}
      \item{Time-dependent technology term: $\mathrm{e}^{\lambda(t-t_0)}$}
      \begin{itemize}
        \item{Represents the irrepressible ``progress'' of technology}
        \item{Is independent of factor of production ($f$)}
      \end{itemize}
      \item{Use for illustration purposes here}
      \item{Example: \citet{Nel:2009wq} use $f$ = $q$}
      \item{Parameters estimated by \texttt{nls} function in \texttt{R}}
    \end{itemize}
\end{frame}

@@

\subsection[Cobb-Douglas]{Evaluation of Cobb-Douglas} %+++++++++++++++++++++++++++++++++++++++

@@

\begin{frame}{Comments on Cobb-Douglas Model Equation} %----------------------------
  \begin{center} \cdeEqn \end{center}
  \begin{itemize}
    \item{Time-dependent technology term: $\mathrm{e}^{\lambda(t-t_0)}$}
    \begin{itemize}
      \item{Represents the irrepressible ``progress'' of technology}
      \item{Is independent of factors of production ($k$, $l$, and $e$)}
    \end{itemize}
    \item{$k$, $l$, and $e$ factors of production ``blend'' with $\alpha + \beta + \gamma = 1$}
    \item{Example: \citet{Froling:2009vo}}
  \end{itemize}
\end{frame}

\begin{frame}{Parameter Estimation for Cobb-Douglas With Energy} %----------------------------
  \begin{center}\cdeEqn\end{center}
  \begin{itemize}
    \item{Objective: determine $\lambda$, $\alpha$, $\beta$, and $\gamma$ that best fits historical data}
    \item{Perform grid search over $\gamma$: Re-estimate $\lambda$, $\alpha$, and $\beta$ at each value of $\gamma$}
    \item{Respect sum constraint: $\alpha + \beta + \gamma = 1$}
    \item{Respect absolute constraints: $0 \leq \alpha, \beta, \gamma \leq 1$} (not yet)
    \item{Parameters estimated by \texttt{nls} function in \texttt{R}}
  \end{itemize}
\end{frame}

@@

\subsection[CES]{Evaluation of CES} %+++++++++++++++++++++++++++++++++++++++

@@

\begin{frame}{Comments on CES Model Equation} %----------------------------
  \begin{center} \small \cesEqn \normalsize \end{center}
  \begin{itemize}
    \item{Time-dependent technology term: $\mathrm{e}^{\lambda(t-t_0)}$}
    \begin{itemize}
      \item{Represents the irrepressible ``progress'' of technology}
      \item{Is independent of factors of production ($k$, $l$, and $e$)}
    \end{itemize}
    \item{$\delta$: split between ($kl$) and ($e$)}
    \item{$\sigma$: elasticity of substitution ($0.0 \leq \sigma \leq 1.0$)}
    \begin{itemize}
      \item{$\sigma$ ******* rightarrow******* 0: ($kl$) and ($e$) are complementary}
      \item{$\sigma$ ******* rightarrow******* 1: ($kl$) and ($e$) are substitutes}
    \end{itemize}
    \item{Example: \citet{Stern:2012ey}}
    \item{Parameters estimated by \texttt{cesEst} function in \texttt{R}'s \texttt{micEcon} package}
  \end{itemize}
\end{frame}

@@

\subsection[LINEX]{Evaluation of LINEX} %+++++++++++++++++++++++++++++++++++++++

@@

\begin{frame}{Comments on LINEX Model Equation} %----------------------------
  \begin{center} \linexEqn \end{center}
  \begin{itemize}
    \item{Derived from thermodynamic considerations}
    \item{Foundation is energy ($e$)}
    \item{No time-dependent technology term}
    \item{Ratios among $k$, $l$, and $e$ ``trim'' energy}
    \item{Example: \citet{Warr:2006hc}}
    \item{Parameters estimated by \texttt{nls} function in \texttt{R}}
  \end{itemize}
\end{frame}

@@

<<Coefs_UK_LINEX>>=
linexqUK <- linexData("UK", "Q")
linexxUK <- linexData("UK", "X")
linexuUK <- linexData("UK", "U")
@
\begin{frame}{LINEX $y$ Predictions} %----------------------------
  \begin{columns}[c]
    \column{2.2in}
      \begin{center}
        \linexEqn

        \tiny
          \begin{tabular}{c|c c}
            $e$ & $a_0$ & $c_t$\\
            \hline
            $q$ & \Sexpr{pprint(linexqUK["LINEX", "a_0"])} & \Sexpr{pprint(linexqUK["LINEX", "c_t"])}\\
            \hline
            $x$ & \Sexpr{pprint(linexxUK["LINEX", "a_0"])} & \Sexpr{pprint(linexxUK["LINEX", "c_t"])}\\
            \hline
            $u$ & \Sexpr{pprint(linexuUK["LINEX", "a_0"])} & \Sexpr{pprint(linexuUK["LINEX", "c_t"])}\\
          \end{tabular}
        \normalsize
      \end{center}
    \column{2.2in}
<<Results_Graph_UK_LINEX, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight>>=
createLINEXLatticeGraph("UK", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
  \end{columns}
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{UK}{0.055}
\end{frame}

<<Coefs_ZA_LINEX>>=
linexqZA <- linexData("ZA", "Q")
linexxZA <- linexData("ZA", "X")
@
\begin{frame}{LINEX $y$ Predictions} %----------------------------
  \begin{columns}[c]
    \column{2.2in}
      \begin{center}
        \linexEqn

        \tiny
          \begin{tabular}{c|c c}
            $e$ & $a_0$ & $c_t$\\
            \hline
            $q$ & \Sexpr{pprint(linexqZA["LINEX", "a_0"])} & \Sexpr{pprint(linexqZA["LINEX", "c_t"])}\\
            \hline
            $x$ & \Sexpr{pprint(linexxZA["LINEX", "a_0"])} & \Sexpr{pprint(linexxZA["LINEX", "c_t"])}\\
          \end{tabular}
        \normalsize
      \end{center}
    \column{2.2in}
<<Results_Graph_ZA_LINEX, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight>>=
createLINEXLatticeGraph("ZA", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
  \end{columns}
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{ZA}{0.055}
\end{frame}

<<Coefs_SA_LINEX>>=
linexqSA <- linexData("SA", "Q")
linexxSA <- linexData("SA", "X")
@
\begin{frame}{LINEX $y$ Predictions} %----------------------------
  \begin{columns}[c]
    \column{2.2in}
      \begin{center}
        \linexEqn

        \tiny
          \begin{tabular}{c|c c}
            $e$ & $a_0$ & $c_t$\\
            \hline
            $q$ & \Sexpr{pprint(linexqSA["LINEX", "a_0"])} & \Sexpr{pprint(linexqSA["LINEX", "c_t"])}\\
            \hline
            $x$ & \Sexpr{pprint(linexxSA["LINEX", "a_0"])} & \Sexpr{pprint(linexxSA["LINEX", "c_t"])}\\
          \end{tabular}
        \normalsize
      \end{center}
    \column{2.2in}
<<Results_Graph_SA_LINEX, fig.width=presentationGraph2ColWidth, fig.height=presentationGraph1RowHeight>>=
createLINEXLatticeGraph("SA", textScaling=textScaleForGraphs, keyXLoc=keyXLoc, keyYLoc=keyYLoc)
@
  \end{columns}
  % Including at bottom of frame so as not to push other material downward in the frame.
  \putFlagUpperRight{SA}{0.055}
\end{frame}

\begin{frame}{95\% Confidence Intervals}{LINEX with $q$} %----------------------------
<<LINEXq_Params_Graph, fig.width=presentationGraph1ColWidth*1.8, fig.height=presentationGraph1RowHeight*1.8>>=
createLINEXParamsGraph("Q")
@  
\end{frame}

\begin{frame}{Implications from LINEX} %----------------------------
  \begin{center}\linexEqn\end{center}
  \vspace{\baselineskip}

  \begin{itemize}
    \item{LINEX parameters $a_0$ and $c_t$ are significant (different from 0.0) for most countries}
    \item{Direct energy input to economy ($q$) is important as a LINEX predictor of economic output ($y$)}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Tentative Conclusions} %----------------------------
  \begin{itemize}
  \item{Including energy in Cobb-Douglas and CES production functions is unwarranted \\
        \scriptsize
        Statistical evidence indicates that energy parameters are not significant for most countries \\
        Why? Energy is \emph{always already} embedded in capital stock ($K$) and labor ($L$), so including energy is unnecessary.
        \normalsize}
  \item{Including energy in LINEX production function is required \\
        \scriptsize
        The foundation of LINEX is an energy term.\\
        Statistical evidence indicates that LINEX parameters $a_0$ and $c_t$ are significant for most countries. \\
        LINEX fits overall are worse than Cobb-Douglas or CES.
        \normalsize}
  \vspace{\baselineskip}
  \pause

  \item{Need a modeling method (that is better than LINEX) that captures energy's embeddedness \\
        \scriptsize
        System dynamics modeling?
        \normalsize}
  \end{itemize}
\end{frame}

\begin{frame}{Next Steps (Summer 2013)} %----------------------------
  \begin{itemize}
    \item{Estimate statistical properties for \emph{all} fitting parameters}
    \vspace{\baselineskip}
    \item{Incorporate all constraints for Cobb-Douglas model}
    \vspace{\baselineskip}
    \item{Complete paper for \emph{Energy Economics}}
  \end{itemize}  
\end{frame}
